\propertitle

### 2019-11-16 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-16T18:08:44+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

A natural procedure to compare the informational sufficiency of two
different sets of moments $M'\defd\set{1,\dotsc,m'}$ and
$M''\defd \set{1,\dotsc,m''}$ is as follows:
\begin{enumerate*}[label=(\roman*)]
\item calculate the relative entropies of the measured frequencies $\yf$
  and the sample marginals of the maximum-entropy distributions built with
  the two corresponding sets of constraints;
\item multiply these relative entropies by the number of time bins of the
  recording;
\item take their difference.
\end{enumerate*}
In formulae,
\begin{multline}
  \label{eq:measure_suffic}
T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}
-
T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}\\
{}\equiv
T  \sum_{\ya} \yff_{\ya} \log\frac{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}.
\end{multline}

This is a natural measure because

@@

 This procedure is equivalent to calculating and comparing the
probabilities of our hypotheses about sufficiency, conditional the data, if
their pre-data probability to be equal; so that for example
\begin{equation}
  \label{eq:compare_hypoth_conditional}
  \pf(\text{\small\enquote{first two moments are sufficient}} \| \text{\small data}) \propto
  \pf(\text{\small data} \|
  \text{\small\enquote{first two moments are sufficient}}).
\end{equation}

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-16T19:24:32+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

A natural procedure to compare the informational sufficiency of two
different sets of moments $M'\defd\set{1,\dotsc,m'}$ and
$M''\defd \set{1,\dotsc,m''}$ is as follows:
\begin{enumerate*}[label=(\roman*)]
\item calculate the relative entropies of the measured frequencies $\yf$
  and the sample marginals of the maximum-entropy distributions built with
  the two corresponding sets of constraints;
\item multiply these relative entropies by the number of time bins of the
  recording;
\item take their difference.
\end{enumerate*}
In formulae,
\begin{multline}
  \label{eq:measure_suffic}
  \text{difference in informational sufficiency between $M'$ and $M''$} \defd{}\\
T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}
-
T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')} \equiv{}\\
T  \sum_{\ya} \yff_{\ya} \ln\frac{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}.
\end{multline}
Why is this a natural measure? Let's ask how much more probable is the
sufficiency of one set with respect to the other, conditional on our data
$\yf$:
\begin{equation}
  \label{eq:ratio_prob_suffic}
  \pf(M'' \| \yf)/
  \pf(M' \| \yf).
\end{equation}
Now, the probability of observing activity $\ya$ in the sample at any time
bin is the sample marginal of the maximum-entropy distribution for the full
network, owing to the excheangeability assumption implicit in the
maximum-entropy method:
\begin{equation}
  \label{eq:sample_marginal}
  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\px(\yA \| M).
\end{equation}
The probability of observing one sequence $(\ya_{t})$ with frequencies
$\yf$ is therefore
\begin{equation}
  \label{eq:prob_frequencies}
  \prod_{t=1}^{T} \pxx(\ya_{t} \| M) \equiv
  \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
\end{equation}
The probability of observing the frequencies $\yf$ is obtained multiplying this
by their multiplicity factor, the multinomial coefficient
\begin{equation}
  \label{eq:multiplicity_factor}
  \binom{T}{T\yf} \defd \frac{T!}{\prod_{\ya}(T\,\yff_{\ya})!}
  \approx \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}},
\end{equation}
the last expression coming from Stirling's approximation
\citep[Lemma~2.2]{csiszaretal2004b}.
If we assign equal pre-data probabilities to the two hypotheses $M'$ and $M''$,
each probability in the ratio~\eqref{eq:ratio_prob_suffic} then becomes, by
Bayes's theorem,
\begin{multline}
  \label{eq:prob_proportional_relentropy}
  \pf(M \| \yf) \propto
  \pf(\yf \| M) \times\text{\small const} \propto
  \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
    T\,\yff_{\ya}}
   \approx{}\\
  \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}}
  \times
  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
    T\,\yff_{\ya}}.
\end{multline}
The logarithm of the probability above is easily seen to be the number of
bins $T$ multiplied by relative entropy between the frequency distribution
$\yf$ and the sample marginal of the maximum-entropy distribution.

Thus, the difference~\eqref{eq:measure_suffic} is the logarithm of the
probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
difference~\eqref{eq:measure_suffic} tells us how much more probable is the
set $M''$ to be sufficient than the set $M'$.

@@
